### 2021-05-21 authlist_end_SNWG.tex
End of final reader

% Author list file generated with: mkauthlist 0+unknown 
% mkauthlist DES-2021-0642_author_list_end_SNWG.csv authlist_end_SNWG.tex -a order.txt -j mnras -sn -f

\author[DES Collaboration]{
\parbox{\textwidth}{
\Large
P.~Wiseman,$^{1}$
M.~Sullivan,$^{1}$
M.~Smith,$^{1}$
C.~Frohmaier,$^{2}$
M.~Vincenzi,$^{2}$
O.~Graur,$^{2}$
B.~Popovic,$^{3}$
P.~Armstrong,$^{4}$
J.~Asorey,$^{5}$
D.~Brout,$^{6,7}$
D.~Carollo,$^{6,8}$
L.~Galbany,$^{9}$
S.~R.~Hinton,$^{10}$
L.~Kelsey,$^{1}$
G.~F.~Lewis,$^{11}$
C.~Lidman,$^{12,4}$
A.~M\"oller,$^{13}$
B.~Rose,$^{3}$
and B.~E.~Tucker$^{4}$
\begin{center} (DES Collaboration) \end{center}
}
\vspace{0.4cm}
\\
\parbox{\textwidth}{
%\scriptsize
$^{1}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK\\
$^{2}$ Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK\\
$^{3}$ Department of Physics, Duke University Durham, NC 27708, USA\\
$^{4}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia\\
$^{5}$ Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain\\
$^{6}$ \\
$^{7}$ Center for Astrophysics $\vert$ Harvard \& Smithsonian, 60 Garden Street, Cambridge, MA 02138, USA\\
$^{8}$ INAF, Astrophysical Observatory of Turin, I-10025 Pino Torinese, Italy\\
$^{9}$ PITT PACC, Department of Physics and Astronomy, University of Pittsburgh, Pittsburgh, PA 15260, USA\\
$^{10}$ School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia\\
$^{11}$ Sydney Institute for Astronomy, School of Physics, A28, The University of Sydney, NSW 2006, Australia\\
$^{12}$ Centre for Gravitational Astrophysics, College of Science, The Australian National University, ACT 2601, Australia\\
$^{13}$ Universit\{'e} Clermont Auvergne, CNRS/IN2P3, LPC, F-63000 Clermont-Ferrand, France\\
}
}