### 2021-05-21 authlist_end_FR.tex
updating bib

% Author list file generated with: mkauthlist 0+unknown 
% mkauthlist DES-2021-0642_author_list_end_FR.csv authlist_end_FR.tex -j mnras -a order.txt -f

\author[DES Collaboration]{
\parbox{\textwidth}{
\Large
P.~Wiseman,$^{1}$
M.~Sullivan,$^{1}$
M.~Smith,$^{2,1}$
C.~Frohmaier,$^{3,1}$
M.~Vincenzi,$^{3}$
O.~Graur,$^{3}$
B.~Popovic,$^{4}$
P.~Armstrong,$^{5}$
D.~Brout,$^{6,7}$
T.~M.~Davis,$^{8}$
L.~Galbany,$^{9}$
S.~R.~Hinton,$^{8}$
L.~Kelsey,$^{1}$
R.~Kessler,$^{10,11}$
C.~Lidman,$^{12,5}$
A.~M\"oller,$^{13}$
R.~C.~Nichol,$^{3}$
B.~Rose,$^{4}$
D.~Scolnic,$^{4}$
M.~Toy,$^{1}$
Z.~Zontou,$^{1}$
J.~Asorey,$^{14}$
D.~Carollo,$^{15}$
K.~Glazebrook,$^{16}$
G.~F.~Lewis,$^{17}$
B.~E.~Tucker,$^{5}$
T.~M.~C.~Abbott,$^{18}$
M.~Aguena,$^{19}$
S.~Allam,$^{20}$
F.~Andrade-Oliveira,$^{21,19}$
J.~Annis,$^{20}$
D.~Bacon,$^{3}$
E.~Bertin,$^{22,23}$
D.~Brooks,$^{24}$
E.~Buckley-Geer,$^{10,20}$
D.~L.~Burke,$^{25,26}$
A.~Carnero~Rosell,$^{27,19,28}$
M.~Carrasco~Kind,$^{29,30}$
J.~Carretero,$^{31}$
M.~Costanzi,$^{32,33,34}$
L.~N.~da Costa,$^{19,35}$
M.~E.~S.~Pereira,$^{36}$
S.~Desai,$^{37}$
H.~T.~Diehl,$^{20}$
P.~Doel,$^{24}$
S.~Everett,$^{38}$
I.~Ferrero,$^{39}$
B.~Flaugher,$^{20}$
P.~Fosalba,$^{40,41}$
J.~Frieman,$^{20,11}$
J.~Garc\'ia-Bellido,$^{42}$
E.~Gaztanaga,$^{40,41}$
T.~Giannantonio,$^{43,44}$
D.~Gruen,$^{45,25,26}$
R.~A.~Gruendl,$^{29,30}$
J.~Gschwend,$^{19,35}$
G.~Gutierrez,$^{20}$
D.~L.~Hollowood,$^{38}$
K.~Honscheid,$^{46,47}$
B.~Hoyle,$^{48,49}$
D.~J.~James,$^{7}$
E.~Krause,$^{50}$
K.~Kuehn,$^{51,52}$
N.~Kuropatkin,$^{20}$
M.~A.~G.~Maia,$^{19,35}$
J.~L.~Marshall,$^{53}$
P.~Martini,$^{46,54,55}$
F.~Menanteau,$^{29,30}$
R.~Miquel,$^{56,31}$
R.~Morgan,$^{57}$
R.~L.~C.~Ogando,$^{19,35}$
A.~Palmese,$^{20,11}$
F.~Paz-Chinch\'{o}n,$^{29,43}$
D.~Petravick,$^{29}$
A.~Pieres,$^{19,35}$
A.~A.~Plazas~Malag\'on,$^{58}$
A.~K.~Romer,$^{59}$
E.~Sanchez,$^{14}$
V.~Scarpine,$^{20}$
M.~Schubnell,$^{36}$
S.~Serrano,$^{40,41}$
I.~Sevilla-Noarbe,$^{14}$
M.~Soares-Santos,$^{36}$
E.~Suchyta,$^{60}$
M.~E.~C.~Swanson,$^{29}$
G.~Tarle,$^{36}$
D.~Thomas,$^{3}$
C.~To,$^{45,25,26}$
T.~N.~Varga,$^{49,61}$
and A.~R.~Walker$^{18}$
\begin{center} (DES Collaboration) \end{center}
}
\vspace{0.4cm}
\\
\parbox{\textwidth}{
%\scriptsize
$^{1}$ School of Physics and Astronomy, University of Southampton,  Southampton, SO17 1BJ, UK\\
$^{2}$ Universit\'e de Lyon 1, CNRS/IN2P3, Institut de Physique des Deux Infinis, 69622 Villeurbanne Cedex, France\\
$^{3}$ Institute of Cosmology and Gravitation, University of Portsmouth, Portsmouth, PO1 3FX, UK\\
$^{4}$ Department of Physics, Duke University Durham, NC 27708, USA\\
$^{5}$ The Research School of Astronomy and Astrophysics, Australian National University, ACT 2601, Australia\\
$^{6}$ NASA Einstein Fellow\\
$^{7}$ Center for Astrophysics $\vert$ Harvard \& Smithsonian, 60 Garden Street, Cambridge, MA 02138, USA\\
$^{8}$ School of Mathematics and Physics, University of Queensland,  Brisbane, QLD 4072, Australia\\
$^{9}$ PITT PACC, Department of Physics and Astronomy, University of Pittsburgh, Pittsburgh, PA 15260, USA\\
$^{10}$ Department of Astronomy and Astrophysics, University of Chicago, Chicago, IL 60637, USA\\
$^{11}$ Kavli Institute for Cosmological Physics, University of Chicago, Chicago, IL 60637, USA\\
$^{12}$ Centre for Gravitational Astrophysics, College of Science, The Australian National University, ACT 2601, Australia\\
$^{13}$ Universit\{'e} Clermont Auvergne, CNRS/IN2P3, LPC, F-63000 Clermont-Ferrand, France\\
$^{14}$ Centro de Investigaciones Energ\'eticas, Medioambientales y Tecnol\'ogicas (CIEMAT), Madrid, Spain\\
$^{15}$ INAF, Osservatorio Astronomico di Trieste, I-34143 Trieste, Italy\\
$^{16}$ Centre for Astrophysics \& Supercomputing, Swinburne University of Technology, Victoria 3122, Australia\\
$^{17}$ Sydney Institute for Astronomy, School of Physics, A28, The University of Sydney, NSW 2006, Australia\\
$^{18}$ Cerro Tololo Inter-American Observatory, NSF's National Optical-Infrared Astronomy Research Laboratory, Casilla 603, La Serena, Chile\\
$^{19}$ Laborat\'orio Interinstitucional de e-Astronomia - LIneA, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil\\
$^{20}$ Fermi National Accelerator Laboratory, P. O. Box 500, Batavia, IL 60510, USA\\
$^{21}$ Instituto de F\'{i}sica Te\'orica, Universidade Estadual Paulista, S\~ao Paulo, Brazil\\
$^{22}$ CNRS, UMR 7095, Institut d'Astrophysique de Paris, F-75014, Paris, France\\
$^{23}$ Sorbonne Universit\'es, UPMC Univ Paris 06, UMR 7095, Institut d'Astrophysique de Paris, F-75014, Paris, France\\
$^{24}$ Department of Physics \& Astronomy, University College London, Gower Street, London, WC1E 6BT, UK\\
$^{25}$ Kavli Institute for Particle Astrophysics \& Cosmology, P. O. Box 2450, Stanford University, Stanford, CA 94305, USA\\
$^{26}$ SLAC National Accelerator Laboratory, Menlo Park, CA 94025, USA\\
$^{27}$ Instituto de Astrofisica de Canarias, E-38205 La Laguna, Tenerife, Spain\\
$^{28}$ Universidad de La Laguna, Dpto. Astrofísica, E-38206 La Laguna, Tenerife, Spain\\
$^{29}$ Center for Astrophysical Surveys, National Center for Supercomputing Applications, 1205 West Clark St., Urbana, IL 61801, USA\\
$^{30}$ Department of Astronomy, University of Illinois at Urbana-Champaign, 1002 W. Green Street, Urbana, IL 61801, USA\\
$^{31}$ Institut de F\'{\i}sica d'Altes Energies (IFAE), The Barcelona Institute of Science and Technology, Campus UAB, 08193 Bellaterra (Barcelona) Spain\\
$^{32}$ Astronomy Unit, Department of Physics, University of Trieste, via Tiepolo 11, I-34131 Trieste, Italy\\
$^{33}$ INAF-Osservatorio Astronomico di Trieste, via G. B. Tiepolo 11, I-34143 Trieste, Italy\\
$^{34}$ Institute for Fundamental Physics of the Universe, Via Beirut 2, 34014 Trieste, Italy\\
$^{35}$ Observat\'orio Nacional, Rua Gal. Jos\'e Cristino 77, Rio de Janeiro, RJ - 20921-400, Brazil\\
$^{36}$ Department of Physics, University of Michigan, Ann Arbor, MI 48109, USA\\
$^{37}$ Department of Physics, IIT Hyderabad, Kandi, Telangana 502285, India\\
$^{38}$ Santa Cruz Institute for Particle Physics, Santa Cruz, CA 95064, USA\\
$^{39}$ Institute of Theoretical Astrophysics, University of Oslo. P.O. Box 1029 Blindern, NO-0315 Oslo, Norway\\
$^{40}$ Institut d'Estudis Espacials de Catalunya (IEEC), 08034 Barcelona, Spain\\
$^{41}$ Institute of Space Sciences (ICE, CSIC),  Campus UAB, Carrer de Can Magrans, s/n,  08193 Barcelona, Spain\\
$^{42}$ Instituto de Fisica Teorica UAM/CSIC, Universidad Autonoma de Madrid, 28049 Madrid, Spain\\
$^{43}$ Institute of Astronomy, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK\\
$^{44}$ Kavli Institute for Cosmology, University of Cambridge, Madingley Road, Cambridge CB3 0HA, UK\\
$^{45}$ Department of Physics, Stanford University, 382 Via Pueblo Mall, Stanford, CA 94305, USA\\
$^{46}$ Center for Cosmology and Astro-Particle Physics, The Ohio State University, Columbus, OH 43210, USA\\
$^{47}$ Department of Physics, The Ohio State University, Columbus, OH 43210, USA\\
$^{48}$ Faculty of Physics, Ludwig-Maximilians-Universit\"at, Scheinerstr. 1, 81679 Munich, Germany\\
$^{49}$ Max Planck Institute for Extraterrestrial Physics, Giessenbachstrasse, 85748 Garching, Germany\\
$^{50}$ Department of Astronomy/Steward Observatory, University of Arizona, 933 North Cherry Avenue, Tucson, AZ 85721-0065, USA\\
$^{51}$ Australian Astronomical Optics, Macquarie University, North Ryde, NSW 2113, Australia\\
$^{52}$ Lowell Observatory, 1400 Mars Hill Rd, Flagstaff, AZ 86001, USA\\
$^{53}$ George P. and Cynthia Woods Mitchell Institute for Fundamental Physics and Astronomy, and Department of Physics and Astronomy, Texas A\&M University, College Station, TX 77843,  USA\\
$^{54}$ Department of Astronomy, The Ohio State University, Columbus, OH 43210, USA\\
$^{55}$ Radcliffe Institute for Advanced Study, Harvard University, Cambridge, MA 02138\\
$^{56}$ Instituci\'o Catalana de Recerca i Estudis Avan\c{c}ats, E-08010 Barcelona, Spain\\
$^{57}$ Physics Department, 2320 Chamberlin Hall, University of Wisconsin-Madison, 1150 University Avenue Madison, WI  53706-1390\\
$^{58}$ Department of Astrophysical Sciences, Princeton University, Peyton Hall, Princeton, NJ 08544, USA\\
$^{59}$ Department of Physics and Astronomy, Pevensey Building, University of Sussex, Brighton, BN1 9QH, UK\\
$^{60}$ Computer Science and Mathematics Division, Oak Ridge National Laboratory, Oak Ridge, TN 37831\\
$^{61}$ Universit\"ats-Sternwarte, Fakult\"at f\"ur Physik, Ludwig-Maximilians Universit\"at M\"unchen, Scheinerstr. 1, 81679 M\"unchen, Germany\\
}
}